Finite element analysis of multi-dimensional and simplified models for beams

### 2024-04-18 final_seminar/main.tex
everything except a concludion

            \begin{itemize}
                \item[-] Provides a method to calculate the exact eigenvalues.
                \item[-] Uses a cantilever beam as an example in the article.
                \item[-] Eigenvalues can be numerically obtained to any degree of accuracy.
            \end{itemize}

@@

        \begin{frame}{Eigenvalue problem - Two and Three-dimensional models}
            General eigenvalue problem for the two- and three-dimensional models:

                For the eigenvalue problem, assume that $M_{f}Q^I = 0$ so that 
            \begin{eqnarray}
                    M\ddot{\bar{u}} & = & K\bar{u}.\label{eq:2DFEM:M2}
            \end{eqnarray}

            Not so simple to compute the eigenvalues and eigenfunctions.

            Need to use a numerical method like FEM to approximate the eigenvalues and eigenvectors.

@@

        \begin{frame}{FEM}
        \end{frame}

        \begin{frame}{Accuracy of the Eigenvalues}
            \centering
            \includegraphics[scale=0.3]{Convergence.png} \\ % Adjust scale as needed
            \footnotesize Rate of convergence of the first 20 eigenvalues.

            \centering
            \includegraphics[scale=0.3]{convergence3d.png} \\ % Adjust scale as needed
            \footnotesize Rate of convergence of the first 20 eigenvalues.
        \end{frame}

    \subsection{Validity of the Timoshenko beam model}
        \begin{frame}{Validity of the Timoshenko beam model}
            \begin{center}
                \begin{minipage}[b]{0.45\linewidth}
                    \centering
                    \begin{tikzpicture}[scale=0.7]
                        \draw[line width = 0.4mm] (0,0.2) -- (6,0.2);
                                    \draw[line width = 0.4mm] (0,-0.2) -- (6,-0.2);
                                    \draw[line width = 0.4mm] (6,-0.2) -- (6,0.2);

@@

        \begin{frame}{Comparison of two and three-dimensional models}
            \centering
            \footnotesize
            \renewcommand{\arraystretch}{0.9}
            Results from [LVV09] and results obtained. $0^a$ indicates a 0 due to rounding. $\alpha = 1200$

            \begin{tabular}{cccc|cccc}
                \hline
                & \multicolumn{3}{c}{[LVV09]} & & \multicolumn{3}{c}{Dissertation} \\

@@

        \begin{frame}{Eigenvalue problem - FEM}
            \centering
            \includegraphics[scale=0.45]{Result.png} \\ % Adjust scale as needed
            \footnotesize Rate of convergence of the first 20 eigenvalues.
            \label{fig:conv}
        \end{frame}

@@

    \subsection{Validity of the Two-dimensional beam model}
        \begin{frame}{Validity of the Two-dimensional beam model}
            \begin{center}
                \begin{minipage}[b]{0.45\linewidth}
                    \centering
                    \begin{tikzpicture}[scale=0.6]
                        \draw[line width = 0.3mm] (-0.1,1) -- (6.9,1);
                        \draw[line width = 0.3mm] (-0.1,-1) -- (6.9,-1);
                        \draw[line width = 0.3mm] (6.9,-1) -- (6.9,1);
                        \draw[line width = 0.3mm] (-0.1,-1) -- (-0.1,1);

@@

        \begin{frame}
            \begin{table}[htbp]
                \scalebox{.8}{
                \makebox[\textwidth]{
                    \begin{tabular}{|cc|cc|cc|cc||cc|}
                        \hline
                        \multicolumn{10}{|c|}{Eigenvalues for $b \leq h$} \\
                        \hline
                        \hline
                        i     & {b = h} & i     & {b = 1/2 h} & i     & {b = 1/4 h} & i     & {b = 1/8 h} & j     & 2D \\
                        \hline
                        {2} & 0.12307 & {2} & 0.12234 & {2} & 0.12198 & {3} & 0.12178 & 1     & 0.12151 \\
                        {3} & 3.5773 & {5} & 3.5630 & {6} & 3.5558 & {8} & 3.5519 & 2     & 3.5460 \\
                        \rowcolor{lightgray}{5} & 7.7799 & {6} & 7.7596 & {8} & 7.7471 & {11} & 7.7401 & 3     & 7.7311 \\
                        {8} & 20.334 & {9} & 20.283 & {11} & 20.26 & {14} & 20.247 & 4     & 20.225 \\
                        {10} & 56.247 & {12} & 56.173 & {15} & 56.156 & {22} & 56.142 & 5     & 56.109 \\
                        \rowcolor{lightgray}{11} & 69.197 & {14} & 69.319 & {17} & 69.281 & {24} & 69.238 & 6     & 69.164 \\
                        {14} & 114.03 & {16} & 114.01 & {20} & 114.05 & {29} & 114.06 & 7     & 114.03 \\
                        \rowcolor{lightgray}{17} & 187.01 & {19} & 189.14 & {25} & 189.37 & {36} & 189.34 & 8     & 189.17 \\
                        {18} & 192.21 & {20} & 192.41 & {26} & 192.58 & {37} & 192.63 & 9     & 192.61 \\
                        {21} & 284.76 & {23} & 285.43 & {31} & 285.74 & {42} & 285.84 & 10    & 285.85 \\
                        {23} & 327.57 & {26} & 328.24 & {35} & 328.37 & {46} & 328.40 & 11    & 328.40 \\
                        \rowcolor{lightgray}{25} & 347.77 & {28} & 356.44 & {36} & 357.30 & {50} & 357.33 & 12    & 357.08 \\
                        {27} & 393.69 & {30} & 396.84 & {38} & 397.28 & {53} & 397.37 & 13    & 397.33 \\
                        {30} & 434.46 & {34} & 441.05 & {41} & 441.81 & {57} & 441.99 & 14    & 442.00 \\
                        \rowcolor{lightgray}{31} & 523.65 & {36} & 534.04 & {43} & 534.17 & {63} & 534.03 & 15    & 533.71 \\
                        {34} & 550.51 & {37} & 537.82 & {44} & 538.86 & {64} & 539.06 & 16    & 538.97 \\
                        \rowcolor{lightgray}{37} & 590.86 & {41} & 587.43 & {48} & 594.17 & {65} & 595.58 & 17    & 596.06 \\
                        {39} & 590.86 & {42} & 600.52 & {49} & 602.25 & {67} & 602.69 & 18    & 602.77 \\
                        \rowcolor{lightgray}{42} & 646.21 & {44} & 657.22 & {50} & 658.04 & {71} & 658.06 & 19    & 657.87 \\
                        {44} & 711.07 & {46} & 714.62 & {53} & 717.10 & {73} & 717.51 & 20    & 717.37 \\
                        \hline
                        \hline
                        \multicolumn{2}{|c|}{Max RE: \  2.6069\%} &\multicolumn{2}{c|}{Max RE: \ 1.4469\%}  & \multicolumn{2}{c|}{Max RE: \  0.38192\%}  & \multicolumn{2}{c||}{Max RE: \ 0.22301\%}& \multicolumn{2}{c|}{-} \\
                        \hline
                    \end{tabular}%
                    \label{tab:2v3_1}%
                }}
            \end{table}%
        \end{frame}

>>>>

            \begin{itemize}
                \item[-] Provides a method to calculate the exact eigenvalues.
                \item[-] Uses a cantilever beam as an example in the article.
                \item[-] Eigenvalues can be numerically obtained to specified degree of accuracy.
            \end{itemize}

@@

        \begin{frame}{Finite Element Method}
            It is not trivial to solve the eigenvalue problem for the cantilever two- and three-dimensional beams. The Finite Element Method (FEM) can be used to approximate the eigenvalues and eigenvectors.

            We descrise our beams into a finite number of elements and nodes. For the two-dimensional beam, we can use a grid of rectangular shaped elements. For the three-dimensional beam, we can use a grid of cuboid or brick shaped elements.

            We can also choose our basis functions. In this presentation, bi-cubic and tri-cubic basis function 
            are used. These have the advantage over bi-linear and tri-linear basis function in that they provide faster convergence of the eigenvalues, although they are more complex to program.
        \end{frame}

        \begin{frame}{Eigenvalue problem - Two-dimensional and three-dimensional model}
            \textbf{Eigenvalue problem}

            Find a real number $\lambda$ and a function $\bar{u} \in S^h$ such that
            \begin{eqnarray}
                M\lambda{\bar{u}} & = & K\bar{u}.
            \end{eqnarray}

            \begin{itemize}
                \item $M$ and $K$ are the mass and stiffness matrices. They are derived from the standard Finite Element Matrices.
                \item In this form, the eigenvalue problem is a system of ordinary differential equations.
                \item It is possible to solve this system of equations using standard numerical methods.
            \end{itemize}

@@

        \begin{frame}{Accuracy of the Eigenvalues}
            \centering
            \includegraphics[scale=0.3]{Convergence.png} \\ % Adjust scale as needed
            \footnotesize Rate of convergence of the first 20 eigenvalues for the two-dimensional beam.

            \centering
            \includegraphics[scale=0.3]{convergence3d.png} \\ % Adjust scale as needed
            \footnotesize Rate of convergence of the first 20 eigenvalues for the three-dimensional beam.
        \end{frame}

\section{Validity of the Timoshenko beam model}
    \begin{frame}{Validity of the Timoshenko beam model}
        \begin{center}
            \begin{minipage}[b]{0.45\linewidth}
                \centering
                \begin{tikzpicture}[scale=0.7]
                    \draw[line width = 0.4mm] (0,0.2) -- (6,0.2);
                                \draw[line width = 0.4mm] (0,-0.2) -- (6,-0.2);
                                \draw[line width = 0.4mm] (6,-0.2) -- (6,0.2);

@@

    \begin{frame}{Comparison of two and three-dimensional models}
        \centering
        \footnotesize
        \renewcommand{\arraystretch}{0.9}
        Results from \cite{LVV09} and results obtained. $0^a$ indicates a 0 due to rounding. Height to lenght ratio of 1 to 10.

        \begin{tabular}{cccc|cccc}
            \hline
            & \multicolumn{3}{c}{[LVV09]} & & \multicolumn{3}{c}{Master Dissertation} \\

@@

    % \begin{frame}{Eigenvalue problem - FEM}
    %     \centering
    %     \includegraphics[scale=0.45]{Result.png} \\ % Adjust scale as needed
    %     \footnotesize Rate of convergence of the first 20 eigenvalues.
    %     \label{fig:conv}
    % \end{frame}

@@

\section{Validity of the Two-dimensional beam model}
    \begin{frame}{Validity of the Two-dimensional beam model}
        \begin{center}
            \begin{minipage}[b]{0.45\linewidth}
                \centering
                \begin{tikzpicture}[scale=0.6]
                    \draw[line width = 0.3mm] (-0.1,1) -- (6.9,1);
                    \draw[line width = 0.3mm] (-0.1,-1) -- (6.9,-1);
                    \draw[line width = 0.3mm] (6.9,-1) -- (6.9,1);
                    \draw[line width = 0.3mm] (-0.1,-1) -- (-0.1,1);

@@

    \begin{frame}
        \begin{table}[htbp]
            \scalebox{.8}{
            \makebox[\textwidth]{
                \begin{tabular}{|cc|cc|cc|cc||cc|}
                    \hline
                    \multicolumn{10}{|c|}{{Comparsion of Eigenvalues with $h = 1/5$, with decreasing $b$ and $b < h$.} \\
                    \hline
                    \hline
                    i     & {b = h} & i     & {b = 1/2 h} & i     & {b = 1/4 h} & i     & {b = 1/8 h} & j     & 2D \\
                    \hline
                    {2} & 0.12307 & {2} & 0.12234 & {2} & 0.12198 & {3} & 0.12178 & 1     & 0.12151 \\
                    {3} & 3.5773 & {5} & 3.5630 & {6} & 3.5558 & {8} & 3.5519 & 2     & 3.5460 \\
                    \rowcolor{lightgray}{5} & 7.7799 & {6} & 7.7596 & {8} & 7.7471 & {11} & 7.7401 & 3     & 7.7311 \\
                    {8} & 20.334 & {9} & 20.283 & {11} & 20.26 & {14} & 20.247 & 4     & 20.225 \\
                    {10} & 56.247 & {12} & 56.173 & {15} & 56.156 & {22} & 56.142 & 5     & 56.109 \\
                    \rowcolor{lightgray}{11} & 69.197 & {14} & 69.319 & {17} & 69.281 & {24} & 69.238 & 6     & 69.164 \\
                    {14} & 114.03 & {16} & 114.01 & {20} & 114.05 & {29} & 114.06 & 7     & 114.03 \\
                    \rowcolor{lightgray}{17} & 187.01 & {19} & 189.14 & {25} & 189.37 & {36} & 189.34 & 8     & 189.17 \\
                    {18} & 192.21 & {20} & 192.41 & {26} & 192.58 & {37} & 192.63 & 9     & 192.61 \\
                    {21} & 284.76 & {23} & 285.43 & {31} & 285.74 & {42} & 285.84 & 10    & 285.85 \\
                    {23} & 327.57 & {26} & 328.24 & {35} & 328.37 & {46} & 328.40 & 11    & 328.40 \\
                    \rowcolor{lightgray}{25} & 347.77 & {28} & 356.44 & {36} & 357.30 & {50} & 357.33 & 12    & 357.08 \\
                    {27} & 393.69 & {30} & 396.84 & {38} & 397.28 & {53} & 397.37 & 13    & 397.33 \\
                    {30} & 434.46 & {34} & 441.05 & {41} & 441.81 & {57} & 441.99 & 14    & 442.00 \\
                    \rowcolor{lightgray}{31} & 523.65 & {36} & 534.04 & {43} & 534.17 & {63} & 534.03 & 15    & 533.71 \\
                    {34} & 550.51 & {37} & 537.82 & {44} & 538.86 & {64} & 539.06 & 16    & 538.97 \\
                    \rowcolor{lightgray}{37} & 590.86 & {41} & 587.43 & {48} & 594.17 & {65} & 595.58 & 17    & 596.06 \\
                    {39} & 590.86 & {42} & 600.52 & {49} & 602.25 & {67} & 602.69 & 18    & 602.77 \\
                    \rowcolor{lightgray}{42} & 646.21 & {44} & 657.22 & {50} & 658.04 & {71} & 658.06 & 19    & 657.87 \\
                    {44} & 711.07 & {46} & 714.62 & {53} & 717.10 & {73} & 717.51 & 20    & 717.37 \\
                    \hline
                    \hline
                    \multicolumn{2}{|c|}{Max RE: \  2.6069\%} &\multicolumn{2}{c|}{Max RE: \ 1.4469\%}  & \multicolumn{2}{c|}{Max RE: \  0.38192\%}  & \multicolumn{2}{c||}{Max RE: \ 0.22301\%}& \multicolumn{2}{c|}{-} \\
                    \hline
                \end{tabular}%
                \label{tab:2v3_1}%
            }}
        \end{table}%
    \end{frame}
